---
title: "Lab1-Vignette"
author: "Austin Parker"
date: "`r Sys.Date()`"
output:
rmarkdown::html_vignette:
  toc: true
  toc_float: true
vignette: >
  %\VignetteIndexEntry{Lab1-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab1)
```

# 2. My Descriptive Statistical Fucntions
***
## Sample Mean Function, $\bar x$
This fucntion returns the sample column means, $\bar x_k$, for a given matrix, $\textbf{X}$

```{r}
SampleMean = function(X) {
  X = matrix(0:8,nrow = 3, ncol = 3, byrow = TRUE)       # default matrix "x" for function

  # print(x)

  n = nrow(X)      # define the number of rows in matrix x

  # Sums the columnd and rows
  col.sum = apply(X, 2, sum)
  # row.sum = apply(x, 1, sum)

  # Adds sums to original matrix (not necessary)
  # x <- cbind(x, Rsum = row.sum)
  # x <- rbind(x, Csum = c(col.sum, sum(col.sum)))
  # print(x)

  # Calculates the sample mean for each column
  xbar = 1/n * col.sum
  print(xbar)
}

X = matrix(0:8,nrow = 3, ncol = 3, byrow = TRUE)
  print(X)

SampleMean()
```

## Sample Covariance Matrix, $s_{ik}$
This fucntion returns the sample covariance matrix, $\textbf s_{kk}$,  for a given matrix, $\textbf{X}$


```{r}
SampleCovariance = function(X) {
  X = matrix(c(15,40,37,31,19,47,45,74,16), nrow = 3, ncol = 3, byrow = TRUE)      # create default matrix for function
  
    n=nrow(X)
    p=ncol(X)

    sums=apply(X, MARGIN = 2, FUN=sum) #sum of measurements for each variable
    xbar=(1/n)*sums

    s <- matrix(NA, nrow = p, ncol = p)  # create matrix with same dimensions as input matrix
    v <- vector(mode = "numeric", length = n)

    for (i in 1:p) {
      for (k in 1:p) {
        for (j in 1:n) {
          v[j] <- (X[j,i] - xbar[i]) * (X[j,k] - xbar[k])
        }
    
        s[i,k] <- 1/(n-1) * sum(v)
      }
  
    
    }
    print(s)
}

X = matrix(c(15,40,37,31,19,47,45,74,16), nrow = 3, ncol = 3, byrow = TRUE)
  print(X)

SampleCovariance()


cov(X)
```

## Sample Correlation Matrix, $r_{ik}$
This fucntion returns the sample correlation matrix, $\textbf r_{kk}$, for a given matrix, $\textbf{X}$


```{r}
SampleCorrelation = function(X) {
  X = matrix(c(15,40,37,31,19,47,45,74,16), nrow = 3, ncol = 3, byrow = TRUE)

    n=nrow(X)
    p=ncol(X)
  
    sums=apply(X, MARGIN = 2, FUN=sum) #sum of measurements for each variable
    xbar=(1/n)*sums
  
    s <- matrix(NA, nrow = p, ncol = p)  # create matrix with same dimensions as input matrix
    r <- matrix(NA, nrow = p, ncol = p)  # create matrix with same dimensions as input matrix
    v <- vector(mode = "numeric", length = n)
  
    for (i in 1:p) {
      for (k in 1:p) {
        for (j in 1:n) {
          v[j] <- (X[j,i] - xbar[i]) * (X[j,k] - xbar[k])
        }
  
        s[i,k] <- 1/(n-1) * sum(v)
      }
    }
  
    for (i in 1:p) {
      for (k in 1:p) {
        r[i,k] <- s[i,k] / (sqrt(s[i,i]) * sqrt(s[k,k]))
        # the code below helps diagnose the problem of inserting the line above from filling in the matrix "r" (k,k) k=2 is need for r_{1,2}, but won't be defined for another 3 loops
        # print(c(i,k))
        # print(r)
      }
    }
    print(r)
}

X = matrix(c(15,40,37,31,19,47,45,74,16), nrow = 3, ncol = 3, byrow = TRUE)
  print(X)

SampleCorrelation()

cor(X)
```

# 3. Links
***

[Vignette Help](http://r-pkgs.had.co.nz/vignettes.html)

[My repository](https://github.com/apark83/Lab1)

# Latex Formulae
***

## a. 

Sample mean, $\bar{x}_{k}$

$$  \bar{x}_k = \frac{1}{n} \sum_{j=1}^{n}{x_{jk}} \quad k=1,2,...,p$$

Sample Covariance, $s_{kk}$

$$ s_{ik} = \frac{1}{n} \sum_{j=1}^{n} (x_{ji} - \bar{x}_i)(x_{jk}- \bar{x}_k) \quad i=1,2,...,p, \quad k=1,2,...,p$$

Sample Correlation, $r_{ik}$

$$ r_{ik} = \frac{s_{ik}}{\sqrt{s_{ii}}\sqrt{s_{kk}}} = \frac{ \sum_{j=1}^{n} (x_{ji} - \bar{x}_i)(x_{jk}- \bar{x}_k)} { \sqrt{\sum_{j=1}^{n} (x_{ji} - \bar{x}_i)^2} \sqrt{\sum_{j=1}^{n} (x_{jk} - \bar{x}_k)^2}} $$

## b. Matrix Operations

Sample Mean, $\bar {\textbf x}$

$$ \bar{\textbf{x}} = \frac{1}{n} \textbf{X}'\textbf{1}$$

Sample Correlation, $\textbf S$

$$ \textbf{S} = \frac{1}{n-1} \textbf{X}' \big( \textbf{1} - \frac{1}{n} \textbf{11}' \big) \textbf{X}$$

Sample Correlation Matrix, $\textbf{R} = \textbf{D}^{-1/2} \textbf{S} \textbf{D}^{-1/2}$

$$
\textbf{D}^{1/2}_{(p \times p)} = 
\begin{bmatrix}
  \sqrt{s_{11}} & 0 & \ldots & 0 \\
  0 & \sqrt{s_{22}} & \ldots & 0 \\
  \vdots & \vdots & \ddots & \vdots \\
  0 & 0 & \ldots & \sqrt{s_{pp}}
\end{bmatrix}
$$
Then
$$
\textbf{D}^{-1/2}_{(p \times p)} = 
\begin{bmatrix}
  \frac{1}{\sqrt{s_{11}}} & 0 & \ldots & 0 \\
  0 & \frac{1}{\sqrt{s_{22}}} & \ldots & 0 \\
  \vdots & \vdots & \ddots & \vdots \\
  0 & 0 & \ldots & \frac{1}{\sqrt{s_{pp}}}
\end{bmatrix}
$$
Since
$$
\textbf{S} = 
\begin{bmatrix}
  s_{11} & s_{12} & \ldots & s_{1p} \\
  \vdots & \vdots & \ddots & \vdots \\
  s_{1p} & s_{2p} & \ldots & s_{pp}
\end{bmatrix}
$$

and
$$
\textbf{R} = 
\begin{bmatrix}
  \frac{s_{11}}{\sqrt{s_{11}}\sqrt{s_{11}}} & \frac{s_{12}}{\sqrt{s_{11}}\sqrt{s_{22}}} & \ldots & \frac{s_{1p}}{\sqrt{s_{11}}\sqrt{s_{pp}}} \\
  \vdots & \vdots & \ddots & \vdots \\
 \frac{s_{1p}}{\sqrt{s_{11}}\sqrt{s_{pp}}} & \frac{s_{2p}}{\sqrt{s_{22}}\sqrt{s_{pp}}} & \ldots & \frac{s_{pp}}{\sqrt{s_{pp}}\sqrt{s_{pp}}}
\end{bmatrix}
=
\begin{bmatrix}
  1 & r_{12} & \ldots & r_{1p} \\
  \vdots & \vdots & \ddots & \vdots \\
  r_{1p} & r_{2p} & \ldots & 1
\end{bmatrix}
$$

we have Sample Correlation Simplified, $\textbf{R}$

$$
\textbf{R} = \textbf{D}^{-1/2} \textbf{S} \textbf{D}^{-1/2}
$$

# 5. Table 1.2 Calculations
***
Complete using my functions

## Load Table

```{r}
table1.2 <- read.table("/Users/austinparker/Documents/Classes/Math 4793/Labs/Lab 1/T1-2.DAT")
names(table1.2) <- c("Density", "Machine Direction", "Cross Direction")
table1.2
```


## Sample Means

```{r}
SampleMean(table1.2)
```

## Sample Covariance

```{r}
SampleCovariance(table1.2)

cov(table1.2)
```

## Sample Correlation

```{r}
SampleCorrelation(table1.2)

cor(table1.2)
```

# 7. Assessment of this course
***
1. Clicker quiz each class and lab (Total 10%)
1. 4 assignments (20% Total)
1. Laboratories (10% Total)
1. 2 exams (10% each, Total 20%)
1. 2(4793) projects (10%)
1. 1 final (30%)
